---
title: "Clustering Analysis Comparison"
author: "Lucy Michaels"
format:
  html:
    code-fold: false
execute:
  echo: true
  warning: false
  message: false
  # Ensure figures are embedded as images:
  fig-format: png
  dpi: 144
editor: visual
---

# Data Preparation

```{r}
R.version.string
```

```{r setup}
library(tidyverse)
library(cluster)
library(manydist)
library(clustrd)
library(arrow)
library(fpc)
library(factoextra)
library(janitor)
library(clusterSim)
library(umap)
library(plotly)
```

```{r}
# Load parquet
df_all <- read_parquet("C:/Users/lucyq/Dropbox/AMDP/THESIS/RAnalysis/016_data_ready.parquet")
```

```{r load_df}
# Drop columns
drop_cols <- c(
    "Protocollo", "severity_max"
)

df <- df_all[, setdiff(names(df_all), drop_cols)]
```

We currently have 8 features and 14849 datapoints.\

```{r}
sapply(df, class)
```

```{r}
# Convert to factors
df$natural_light <- factor(df$natural_light, ordered = FALSE)
df <- df %>% mutate(across(where(is.character), as.factor))

# Integer categories to factors
int_as_cats <- c("natural_light", "road_conditions_bin", "time_segment")

df <- df %>% mutate(across(any_of(int_as_cats), as.factor))

# Ensure numeric
df <- df %>% mutate(across(where(is.numeric), as.numeric))
```

# Method 1: ClusPCAmix

## Analysis ClusPCAmix

### Test k = 7

```{r}
# Try 5 dimensions for the internal PCA mix space
n_clusters <- 7
ndim_try <- 5
results_clus_k7d5 <- cluspcamix(df, n_clusters, ndim_try,
    method = "mixedRKM",
    center = TRUE,
    scale = TRUE,
    alpha = NULL,
    rotation = "none",
    nstart = 50,
    seed = c(1234, 2025, 1977)
)

print(results_clus_k7d5)
```

### Test k = 6

```{r}
# Try 5 dimensions for the internal PCA mix space
n_clusters <- 6
ndim_try <- 5
results_clus_k6d5 <- cluspcamix(df, n_clusters, ndim_try,
    method = c("mixedRKM"),
    center = TRUE,
    scale = TRUE,
    alpha = NULL,
    rotation = "none",
    nstart = 50,
    seed = c(1234, 2025, 1977)
)

print(results_clus_k6d5)
```

### Test k = 5

```{r}
# Try 5 dimensions for the internal PCA mix space
n_clusters <- 5
ndim_try <- 5
results_clus_k5d5 <- cluspcamix(df, n_clusters, ndim_try,
    method = c("mixedRKM"),
    center = TRUE,
    scale = TRUE,
    alpha = NULL,
    rotation = "none",
    nstart = 50,
    seed = c(1234, 2025, 1977)
)

print(results_clus_k5d5)
```

###Test k = 5, dim = 4

```{r}
# Try 5 dimensions for the internal PCA mix space
n_clusters <- 5
ndim_try <- 4
results_clus_k5d4 <- cluspcamix(df, n_clusters, ndim_try,
    method = c("mixedRKM"),
    center = TRUE,
    scale = TRUE,
    alpha = NULL,
    rotation = "none",
    nstart = 50,
    seed = c(1234, 2025, 1977)
)

print(results_clus_k5d4)
```

## Results ClusPCAmix

### Results ClusPCAmix k = 7

```{r}
# Silhouette width (higher = better separation)
sil_clus_k7d5 <- silhouette(results_clus_k7d5$cluster, dist(results_clus_k7d5$obscoord))

mean(sil_clus_k7d5[, 3])

fviz_silhouette(sil_clus_k7d5)

summary(sil_clus_k7d5)

```

```{r}
stats_clus_k7d5 <- cluster.stats(dist(results_clus_k7d5$obscoord), results_clus_k7d5$cluster)

# Calculate Davies-Bouldin Index
db_clus_k7d5 <- index.DB(x = results_clus_k7d5$obscoord, 
                          cl = results_clus_k7d5$cluster)$DB

cat(
    "k=7, dim = 5: Silhouette =", mean(sil_clus_k7d5[, 3]),
    ", CH =", stats_clus_k7d5$ch,
    ", Dunn =", stats_clus_k7d5$dunn, 
    ", DB =", db_clus_k7d5, "\n"
)
```

### Results ClusPCAmix k = 6

```{r}
# Silhouette width (higher = better separation)
sil_clus_k6d5 <- silhouette(results_clus_k6d5$cluster, dist(results_clus_k6d5$obscoord))
mean(sil_clus_k6d5[, 3])
fviz_silhouette(sil_clus_k6d5)
summary(sil_clus_k6d5)
```

```{r}
stats_clus_k6d5 <- cluster.stats(dist(results_clus_k6d5$obscoord), results_clus_k6d5$cluster)

# Calculate Davies-Bouldin Index
db_clus_k6d5 <- index.DB(x = results_clus_k6d5$obscoord, 
                          cl = results_clus_k6d5$cluster)$DB

cat(
    "k=6, dim = 5: Silhouette =", mean(sil_clus_k6d5[, 3]),
    ", CH =", stats_clus_k6d5$ch,
    ", Dunn =", stats_clus_k6d5$dunn, 
    ", DB =", db_clus_k6d5, "\n"
)
```

### Results ClusPCAmix k = 5

```{r}
# Silhouette width (higher = better separation)
sil_clus_k5d5 <- silhouette(results_clus_k5d5$cluster, dist(results_clus_k5d5$obscoord))
mean(sil_clus_k5d5[, 3])
fviz_silhouette(sil_clus_k5d5)
summary(sil_clus_k5d5)
```

```{r}
stats_clus_k5d5 <- cluster.stats(dist(results_clus_k5d5$obscoord), results_clus_k5d5$cluster)

# Calculate Davies-Bouldin Index
db_clus_k5d5 <- index.DB(x = results_clus_k5d5$obscoord, 
                          cl = results_clus_k5d5$cluster)$DB

cat(
    "k=5, dim = 5: Silhouette =", mean(sil_clus_k5d5[, 3]),
    ", CH =", stats_clus_k5d5$ch,
    ", Dunn =", stats_clus_k5d5$dunn, 
    ", DB =", db_clus_k5d5,"\n"
)
```

### Test k = 5 but dim = 4

```{r}
# Silhouette width (higher = better separation)
sil_clus_k5d4 <- silhouette(results_clus_k5d4$cluster, dist(results_clus_k5d4$obscoord))
mean(sil_clus_k5d4[, 3])
fviz_silhouette(sil_clus_k5d4)
summary(sil_clus_k5d4)
```

```{r}
stats_clus_k5d4 <- cluster.stats(dist(results_clus_k5d4$obscoord), results_clus_k5d4$cluster)

# Calculate Davies-Bouldin Index
db_clus_k5d4 <- index.DB(x = results_clus_k5d4$obscoord, 
                          cl = results_clus_k5d4$cluster)$DB

cat(
    "k=5, dim = 4: Silhouette =", mean(sil_clus_k5d4[, 3]),
    ", CH =", stats_clus_k5d4$ch,
    ", Dunn =", stats_clus_k5d4$dunn, 
    ", DB =", db_clus_k5d4,"\n"
)
```

# TRY NEW VALUES OF DIMENSIONS

# Test k = 5, d = 3

```{r}
# Try 3 dimensions for the internal PCA mix space
n_clusters <- 5
ndim_try <- 3
results_clus_k5d3 <- cluspcamix(df, n_clusters, ndim_try,
    method = c("mixedRKM"),
    center = TRUE,
    scale = TRUE,
    alpha = NULL,
    rotation = "none",
    nstart = 50,
    seed = c(1234, 2025, 1977)
)

print(results_clus_k5d3)
```

```{r}
# Silhouette width (higher = better separation)
sil_clus_k5d3 <- silhouette(results_clus_k5d3$cluster, dist(results_clus_k5d3$obscoord))
mean(sil_clus_k5d3[, 3])
fviz_silhouette(sil_clus_k5d3)
summary(sil_clus_k5d3)
```

```{r}
stats_clus_k5d3 <- cluster.stats(dist(results_clus_k5d3$obscoord), results_clus_k5d3$cluster)

# Calculate Davies-Bouldin Index
db_clus_k5d3 <- index.DB(x = results_clus_k5d3$obscoord, 
                          cl = results_clus_k5d3$cluster)$DB

cat(
    "k=5, dim = 3: Silhouette =", mean(sil_clus_k5d3[, 3]),
    ", CH =", stats_clus_k5d3$ch,
    ", Dunn =", stats_clus_k5d3$dunn, 
    ", DB =", db_clus_k5d3,"\n"
)
```

## Summary of Stats when k = 5, dim = 3, 4, 5

```{r}
# Summary comparison
comparison <- tibble(
    d = c(3, 4, 5),
    Silhouette = c(mean(sil_clus_k5d3[, 3]), mean(sil_clus_k5d4[, 3]), mean(sil_clus_k5d5[, 3])),
    CH = c(stats_clus_k5d3$ch, stats_clus_k5d4$ch, stats_clus_k5d5$ch),
    Dunn = c(stats_clus_k5d3$dunn, stats_clus_k5d4$dunn, stats_clus_k5d5$dunn), 
    DB = c(db_clus_k5d3, db_clus_k5d4, db_clus_k5d5)
)
print(comparison)
```

# BEST OPTION: k = 5, d = 4

```{r}
library(umap)
library(plotly)

# Create 3D visualization projection
umap3D_k5d4 <- umap(results_clus_k5d4$obscoord, n_components = 3)

# Plot in 3D
plot_ly(x = umap3D_k5d4$layout[,1],
        y = umap3D_k5d4$layout[,2], 
        z = umap3D_k5d4$layout[,3],
        color = as.factor(results_clus_k5d4$cluster),  # convert to factor
        type = "scatter3d",
        mode = "markers") %>%
  layout(scene = list(
    xaxis = list(title = "UMAP 1"),
    yaxis = list(title = "UMAP 2"),
    zaxis = list(title = "UMAP 3")
  ))
```

# EXPERIMENTATION k = 5, d = 4, varying alpha

#### results_clus_k5d4 already exists

```{r}
k5d4_RKM_none_alpha_null <- results_clus_k5d4
```

```{r}
# Try 
k5d4_none_alpha_1 <- cluspcamix(df, 5, 4,
    center = TRUE,
    scale = TRUE,
    alpha = 1,
    rotation = "none",
    nstart = 50,
    seed = c(1234, 2025, 1977)
)
```

```{r}
# Try 
k5d4_none_alpha_0_5 <- cluspcamix(df, 5, 4,
    center = TRUE,
    scale = TRUE,
    alpha = 0.5,
    rotation = "none",
    nstart = 50,
    seed = c(1234, 2025, 1977)
)
```

```{r}
# Try 
k5d4_RKM_promax_alpha_null <- cluspcamix(df, 5, 4,
    method = c("mixedRKM"),
    center = TRUE,
    scale = TRUE,
    alpha = NULL,
    rotation = "promax",
    nstart = 50,
    seed = c(1234, 2025, 1977)
)
```

```{r}
# Try 
k5d4_FKM_none_alpha_null <- cluspcamix(df, 5, 4,
    method = c("mixedFKM"),
    center = TRUE,
    scale = TRUE,
    alpha = NULL,
    rotation = "none",
    nstart = 50,
    seed = c(1234, 2025, 1977)
)
```

```{r}
# Try 
k5d4_FKM_varimax_alpha_null <- cluspcamix(df, 5, 4,
    method = c("mixedFKM"),
    center = TRUE,
    scale = TRUE,
    alpha = NULL,
    rotation = "varimax",
    nstart = 50,
    seed = c(1234, 2025, 1977)
)
```

```{r}
# Try 
k5d4_FKM_promax_alpha_null <- cluspcamix(df, 5, 4,
    method = c("mixedFKM"),
    center = TRUE,
    scale = TRUE,
    alpha = NULL,
    rotation = "promax",
    nstart = 50,
    seed = c(1234, 2025, 1977)
)
```

```{r}
# Silhouette width (higher = better separation)
sil_clus_k5d3 <- silhouette(results_clus_k5d3$cluster, dist(results_clus_k5d3$obscoord))
mean(sil_clus_k5d3[, 3])
fviz_silhouette(sil_clus_k5d3)
summary(sil_clus_k5d3)
```

```{r}
stats_clus_k5d3 <- cluster.stats(dist(results_clus_k5d3$obscoord), results_clus_k5d3$cluster)

# Calculate Davies-Bouldin Index
db_clus_k5d3 <- index.DB(x = results_clus_k5d3$obscoord, 
                          cl = results_clus_k5d3$cluster)$DB

cat(
    "k=5, dim = 3: Silhouette =", mean(sil_clus_k5d3[, 3]),
    ", CH =", stats_clus_k5d3$ch,
    ", Dunn =", stats_clus_k5d3$dunn, 
    ", DB =", db_clus_k5d3,"\n"
)
```

## Summary of Stats when dim = 5

```{r}
# Summary comparison
comparison <- tibble(
    k = c(5, 6, 7),
    Silhouette = c(mean(sil_clus_k5d5[, 3]), mean(sil_clus_k6d5[, 3]), mean(sil_clus_k7d5[, 3])),
    CH = c(stats_clus_k5d5$ch, stats_clus_k6d5$ch, stats_clus_k7d5$ch),
    Dunn = c(stats_clus_k5d5$dunn, stats_clus_k6d5$dunn, stats_clus_k7d5$dunn), 
    DB = c(db_clus_k5d5, db_clus_k6d5, db_clus_k7d5)
)
print(comparison)
```

```{r}
# Add cluster assignments to your original data
data_clustered <- results_clus_k5d4$odata %>%
  mutate(cluster = as.factor(results_clus_k5d4$cluster))

# 1. NUMERICAL FEATURES - Summary statistics by cluster
numerical_summary <- data_clustered %>%
  group_by(cluster) %>%
  summarise(
    n = n(),
    temp_mean = mean(temp_celsius),
    temp_sd = sd(temp_celsius),
    humidity_mean = mean(humidity_percent),
    humidity_sd = sd(humidity_percent),
    precip_mean = mean(precipitation_winsorized),
    precip_sd = sd(precipitation_winsorized),
    wind_mean = mean(wind_gusts_winsorized),
    wind_sd = sd(wind_gusts_winsorized),
    dry_spell_mean = mean(dry_spells_winsorized),
    dry_spell_sd = sd(dry_spells_winsorized)
  )

print(numerical_summary)

# 2. CATEGORICAL FEATURES - Frequency distributions
categorical_summary <- data_clustered %>%
  group_by(cluster) %>%
  count(time_segment, natural_light, road_conditions_bin) %>%
  arrange(cluster, desc(n))

# Top categories per cluster
top_categories <- data_clustered %>%
  group_by(cluster) %>%
  summarise(
    top_time = names(sort(table(time_segment), decreasing = TRUE))[1],
    top_light = names(sort(table(natural_light), decreasing = TRUE))[1],
    top_road = names(sort(table(road_conditions_bin), decreasing = TRUE))[1]
  )

print(top_categories)

# 3. LIFT ANALYSIS (like your K-Prototypes output)
calculate_lift <- function(data, var) {
  data %>%
    group_by(cluster, !!sym(var)) %>%
    summarise(n = n(), .groups = "drop") %>%
    group_by(cluster) %>%
    mutate(cluster_prop = n / sum(n)) %>%
    ungroup() %>%
    group_by(!!sym(var)) %>%
    mutate(overall_prop = sum(n) / nrow(data)) %>%
    ungroup() %>%
    mutate(lift = cluster_prop / overall_prop) %>%
    arrange(cluster, desc(lift))
}

time_lift <- calculate_lift(data_clustered, "time_segment")
light_lift <- calculate_lift(data_clustered, "natural_light")
road_lift <- calculate_lift(data_clustered, "road_conditions_bin")

print("Time Segment Lift:")
print(time_lift)
print("Natural Light Lift:")
print(light_lift)
print("Road Conditions Lift:")
print(road_lift)

# 4. VISUALIZE - Boxplots for numerical features
library(ggplot2)
library(patchwork)

p1 <- ggplot(data_clustered, aes(x = cluster, y = temp_celsius, fill = cluster)) +
  geom_boxplot() + theme_minimal() + labs(title = "Temperature by Cluster", y = "Temperature (°C)")

p2 <- ggplot(data_clustered, aes(x = cluster, y = humidity_percent, fill = cluster)) +
  geom_boxplot() + theme_minimal() + labs(title = "Humidity by Cluster", y = "Humidity (%)")

p3 <- ggplot(data_clustered, aes(x = cluster, y = wind_gusts_winsorized, fill = cluster)) +
  geom_boxplot() + theme_minimal() + labs(title = "Wind by Cluster", y = "Wind Gusts (km/h)")

p4 <- ggplot(data_clustered, aes(x = cluster, y = dry_spells_winsorized, fill = cluster)) +
  geom_boxplot() + theme_minimal() + labs(title = "Dry Spells by Cluster", y = "Dry Spell Duration (days)")

(p1 + p2) / (p3 + p4)

# 5. CLUSTER CENTROIDS in original space
cluster_centroids <- data_clustered %>%
  group_by(cluster) %>%
  summarise(across(where(is.numeric), mean))

print(cluster_centroids)

# 6. STATISTICAL SIGNIFICANCE - ANOVA for numerical features
anova_temp <- aov(temp_celsius ~ cluster, data = data_clustered)
anova_humidity <- aov(humidity_percent ~ cluster, data = data_clustered)
anova_wind <- aov(wind_gusts_winsorized ~ cluster, data = data_clustered)
anova_dry <- aov(dry_spells_winsorized ~ cluster, data = data_clustered)
anova_precip <- aov(precipitation_winsorized ~ cluster, data = data_clustered)

cat("\n=== ANOVA Results ===\n")
cat("\nTemperature:\n")
summary(anova_temp)
cat("\nHumidity:\n")
summary(anova_humidity)
cat("\nWind Gusts:\n")
summary(anova_wind)
cat("\nDry Spells:\n")
summary(anova_dry)
cat("\nPrecipitation:\n")
summary(anova_precip)

# 7. CHI-SQUARE tests for categorical features
cat("\n=== Chi-Square Results ===\n")
cat("\nTime Segment:\n")
print(chisq.test(table(data_clustered$cluster, data_clustered$time_segment)))
cat("\nNatural Light:\n")
print(chisq.test(table(data_clustered$cluster, data_clustered$natural_light)))
cat("\nRoad Conditions:\n")
print(chisq.test(table(data_clustered$cluster, data_clustered$road_conditions_bin)))

# 8. EFFECT SIZE for numerical features (eta-squared)
library(effectsize)
eta_squared(anova_temp)
eta_squared(anova_humidity)
eta_squared(anova_wind)
eta_squared(anova_dry)
eta_squared(anova_precip)
```

Now we try to get profiles of each cluster:

```{r}
# ===== PREPARE DATA =====
# Add cluster assignments to your data
data_clustered <- results_clus_k5d4$odata %>%
  mutate(cluster = as.factor(results_clus_k5d4$cluster))

# Create cluster profiles
cluster_profiles <- data_clustered %>%
  group_by(cluster) %>%
  summarise(n = n()) %>%
  mutate(pct = round(n/sum(n)*100, 1))

# ===== VISUALIZATIONS =====

# 1. Cluster sizes
library(ggplot2)
p1 <- ggplot(cluster_profiles, aes(x = reorder(factor(cluster), -n), y = n)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  geom_text(aes(label = paste0(n, "\n(", pct, "%)")), vjust = -0.3, size = 3) +
  labs(title = "Accident Distribution Across 5 Clusters (ClusPCAmix)",
       x = "Cluster", y = "Number of Accidents") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 0))
print(p1)

# 2. Silhouette plot (you already have this with fviz_silhouette)
fviz_silhouette(sil_clus_k5d4, 
                palette = "jco",
                ggtheme = theme_minimal()) +
  labs(title = "Silhouette Plot (k=5, d=4)")

# 3. Cluster composition by natural light
p2 <- data_clustered %>%
  group_by(cluster, natural_light) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(cluster) %>%
  mutate(pct = n/sum(n)*100) %>%
  ggplot(aes(x = factor(cluster), y = pct, fill = natural_light)) +
  geom_bar(stat = "identity", color = "black") +
  labs(title = "Natural Light Conditions by Cluster",
       x = "Cluster", y = "Percentage", fill = "Light") +
  theme_minimal() +
  scale_fill_manual(values = c("light_day" = "gold", 
                              "light_twilight" = "steelblue", 
                              "light_night" = "navy"))
print(p2)

# 4. Cluster composition by time segment
p3 <- data_clustered %>%
  group_by(cluster, time_segment) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(cluster) %>%
  mutate(pct = n/sum(n)*100) %>%
  ggplot(aes(x = factor(cluster), y = pct, fill = factor(time_segment))) +
  geom_bar(stat = "identity") +
  labs(title = "Time Segment by Cluster",
       x = "Cluster", y = "Percentage", fill = "Time") +
  theme_minimal() +
  scale_fill_brewer(palette = "Set2")
print(p3)

# 5. Cluster composition by road conditions
p4 <- data_clustered %>%
  group_by(cluster, road_conditions_bin) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(cluster) %>%
  mutate(pct = n/sum(n)*100) %>%
  ggplot(aes(x = factor(cluster), y = pct, fill = factor(road_conditions_bin))) +
  geom_bar(stat = "identity") +
  labs(title = "Road Conditions by Cluster",
       x = "Cluster", y = "Percentage", fill = "Road") +
  theme_minimal() +
  scale_fill_brewer(palette = "Dark2")
print(p4)

# 6. Weather profiles by cluster (using pivot_longer for faceted view)
weather_long <- data_clustered %>%
  select(cluster, temp_celsius, precipitation_winsorized, 
         wind_gusts_winsorized, humidity_percent, dry_spells_winsorized) %>%
  pivot_longer(-cluster, names_to = "variable", values_to = "value")

p5 <- ggplot(weather_long, aes(x = factor(cluster), y = value)) +
  geom_boxplot(aes(fill = factor(cluster))) +
  facet_wrap(~variable, scales = "free_y", ncol = 2) +
  labs(title = "Weather Conditions by Cluster",
       x = "Cluster", y = "Value") +
  theme_minimal() +
  theme(legend.position = "none")
print(p5)

# 7. Individual weather variable plots (alternative to faceted view)
library(patchwork)

pw1 <- ggplot(data_clustered, aes(x = cluster, y = temp_celsius, fill = cluster)) +
  geom_boxplot() + 
  theme_minimal() + 
  labs(title = "Temperature", y = "°C") +
  theme(legend.position = "none")

pw2 <- ggplot(data_clustered, aes(x = cluster, y = humidity_percent, fill = cluster)) +
  geom_boxplot() + 
  theme_minimal() + 
  labs(title = "Humidity", y = "%") +
  theme(legend.position = "none")

pw3 <- ggplot(data_clustered, aes(x = cluster, y = wind_gusts_winsorized, fill = cluster)) +
  geom_boxplot() + 
  theme_minimal() + 
  labs(title = "Wind Gusts", y = "km/h") +
  theme(legend.position = "none")

pw4 <- ggplot(data_clustered, aes(x = cluster, y = precipitation_winsorized, fill = cluster)) +
  geom_boxplot() + 
  theme_minimal() + 
  labs(title = "Precipitation", y = "mm") +
  theme(legend.position = "none")

(pw1 + pw2) / (pw3 + pw4) + 
  plot_annotation(title = "Weather Variables by Cluster")

# ===== CLUSTER QUALITY METRICS =====
cat("\n=== CLUSTER QUALITY METRICS (k=5, d=4) ===\n")
cat(sprintf("Average Silhouette Width: %.4f\n", mean(sil_clus_k5d4[, 3])))
cat(sprintf("Calinski-Harabasz Index: %.2f\n", stats_clus_k5d4$ch))
cat(sprintf("Dunn Index: %.4f\n", stats_clus_k5d4$dunn))
cat(sprintf("Davies-Bouldin Index: %.4f\n", db_clus_k5d4))

cat("\nSilhouette Width by Cluster:\n")
sil_summary <- summary(sil_clus_k5d4)
print(sil_summary$clus.avg.widths)

cat("\nCluster with lowest quality: ", 
    which.min(sil_summary$clus.avg.widths), 
    " (Sil=", round(min(sil_summary$clus.avg.widths), 4), ")\n")
cat("Cluster with highest quality: ", 
    which.max(sil_summary$clus.avg.widths), 
    " (Sil=", round(max(sil_summary$clus.avg.widths), 4), ")\n")

# Number of poorly clustered observations
poorly_clustered <- sum(sil_clus_k5d4[, 3] < 0)
cat(sprintf("\nPoorly clustered accidents: %d (%.1f%%)\n", 
            poorly_clustered, poorly_clustered/nrow(data_clustered)*100))

# Cluster sizes
cat("\nCluster Sizes:\n")
print(cluster_profiles)

# ===== SAVE RESULTS =====
cat("\n=== SAVING RESULTS ===\n")

# Save the clustered data
write.csv(data_clustered, "accidents_clustered_k5d4.csv", row.names = FALSE)
cat("✓ Saved: accidents_clustered_k5d4.csv\n")

# Save cluster profiles
write.csv(cluster_profiles, "cluster_profiles_k5d4.csv", row.names = FALSE)
cat("✓ Saved: cluster_profiles_k5d4.csv\n")

# Save clustering info
clustering_info <- list(
  method = "ClusPCAmix_mixedRKM",
  k = 5,
  ndim = 4,
  silhouette_avg = mean(sil_clus_k5d4[, 3]),
  ch_index = stats_clus_k5d4$ch,
  dunn_index = stats_clus_k5d4$dunn,
  db_index = db_clus_k5d4,
  cluster_sizes = table(data_clustered$cluster),
  cluster_assignments = results_clus_k5d4$cluster
)
saveRDS(clustering_info, "clustering_info_k5d4.rds")
cat("✓ Saved: clustering_info_k5d4.rds\n")

# Save the full results object
saveRDS(results_clus_k5d4, "cluspcamix_results_k5d4.rds")
cat("✓ Saved: cluspcamix_results_k5d4.rds\n")

cat("\n=== CLUSTERING VISUALIZATION COMPLETE! ===\n")
```
