---
title: "Clustering Analysis PAM"
author: "Lucy Michaels"
format:
  html:
    code-fold: false
execute:
  echo: true
  warning: false
  message: false
  # Ensure figures are embedded as images:
  fig-format: png
  dpi: 144
editor: visual
---

# Data Preparation

```{r}
R.version.string
```

```{r setup}
library(tidyverse)
library(cluster)
library(manydist)
library(clustrd)
library(arrow)
library(factoextra)
library(dplyr)
library(catdist)

```

```{r}
# Detach manydist to avoid conflicts
detach("package:manydist", unload = TRUE)

# Now reload catdist
library(catdist)
```

```{r}
# Load parquet
df_full <- read_parquet("C:/Users/lucyq/Dropbox/AMDP/THESIS/RAnalysis/016_data_ready.parquet")
```

```{r load_df}
# Drop columns
drop_cols <- c(
    "Protocollo", "severity_max"
)

df <- df_full[, setdiff(names(df_full), drop_cols)]
```

We currently have 8 features and 14849 datapoints.

```{r}
# Convert to factors
df$natural_light <- factor(df$natural_light, ordered = FALSE)
df <- df %>% mutate(across(where(is.character), as.factor))

# Integer categories to factors
int_as_cats <- c("natural_light", "road_conditions_bin", "time_segment")

df <- df %>% mutate(across(any_of(int_as_cats), as.factor))

# Ensure numeric
df <- df %>% mutate(across(where(is.numeric), as.numeric))
```

Now we have 3 features as factors & 5 features as numerical

We use daisy to calculate a dissimilarity matrix, giving the combined, normalized difference between data points accounting for column type, then PAM to test different values of k cluster:

```{r}
# Compute the Gower dissimilarity matrix
dissim <- daisy(df, metric = "gower")

# Test different k values and find optimal
k_range <- 5:10  # Test 4 to 8 clusters
silhouette_scores <- numeric(length(k_range))
pam_models <- list()

cat("Testing different numbers of clusters...\n")
for (i in seq_along(k_range)) {
  k <- k_range[i]
  pam_k <- pam(dissim, k = k, diss = TRUE)
  silhouette_scores[i] <- pam_k$silinfo$avg.width
  pam_models[[i]] <- pam_k
  cat(sprintf("k = %d: Silhouette = %.4f\n", k, silhouette_scores[i]))
}

# Find optimal k (highest silhouette score)
optimal_idx <- which.max(silhouette_scores)
optimal_k <- k_range[optimal_idx]
optimal_silhouette <- silhouette_scores[optimal_idx]

cat(sprintf("\n*** OPTIMAL k = %d (Silhouette = %.4f) ***\n\n", 
            optimal_k, optimal_silhouette))

# Plot silhouette scores
plot(k_range, silhouette_scores, type = "b", 
     xlab = "Number of Clusters (k)", 
     ylab = "Average Silhouette Width",
     main = "Optimal k Selection",
     pch = 19, col = "blue", lwd = 2)
points(optimal_k, optimal_silhouette, col = "red", pch = 19, cex = 2)
text(optimal_k, optimal_silhouette, 
     labels = paste0("k=", optimal_k), 
     pos = 3, col = "red", font = 2)
abline(h = c(0.25, 0.5), lty = 2, col = "gray")
text(max(k_range), 0.25, "Weak", pos = 2, col = "gray")
text(max(k_range), 0.5, "Reasonable", pos = 2, col = "gray")

# Use the optimal clustering
pam_results <- pam_models[[optimal_idx]]

# Add cluster assignment to dataframe
df$cluster <- pam_results$clustering

# Summary
cat("\nCluster Sizes:\n")
print(table(df$cluster))

cat("\nCluster Silhouette Widths:\n")
print(pam_results$silinfo$clus.avg.widths)

# Plot the optimal silhouette
plot(pam_results, which.plots = 2, 
     main = paste0("Silhouette Plot (k=", optimal_k, ")"))
```

```         
```

Instead, we can try the catdist package

```{r}
# Prepare your data (remove any existing cluster column)
df_clean <- df %>% select(-any_of("cluster"))

# Separate categorical and numerical variables
df_categorical <- df_clean %>% 
  select(natural_light, time_segment, road_conditions_bin)

df_numerical <- df_clean %>% 
  select(precipitation_winsorized, temp_celsius, wind_gusts_winsorized, 
         humidity_percent, dry_spells_winsorized)

# 1. Compute catdist for categorical variables
cat("Computing catdist for categorical variables...\n")
dist_cat <- catdist::cdist(df_categorical, method = "tot_var_dist")
D_cat <- dist_cat$distance_mat

# 2. Compute distance for numerical variables (scaled)
cat("Computing distance for numerical variables...\n")
df_numerical_scaled <- scale(df_numerical)  # Standardize
D_num <- dist(df_numerical_scaled, method = "euclidean")  # or "manhattan"
D_num <- as.matrix(D_num)

# 3. Normalize both distance matrices to [0, 1] range
D_cat_norm <- D_cat / max(D_cat)
D_num_norm <- D_num / max(D_num)

# 4. Combine with weights (adjust weights as needed)
weight_cat <- 0.5  # Weight for categorical distance
weight_num <- 0.5  # Weight for numerical distance

D_combined <- weight_cat * D_cat_norm + weight_num * D_num_norm

# 5. Test different k values
k_range <- 5:10
silhouette_scores <- numeric(length(k_range))
pam_models <- list()

cat("\nTesting combined distance with different k values:\n")
for (i in seq_along(k_range)) {
  k <- k_range[i]
  pam_k <- pam(D_combined, k = k, diss = TRUE)
  silhouette_scores[i] <- pam_k$silinfo$avg.width
  pam_models[[i]] <- pam_k
  cat(sprintf("k=%d: Sil=%.4f\n", k, silhouette_scores[i]))
}

# 6. Use optimal k
optimal_k <- k_range[which.max(silhouette_scores)]
optimal_sil <- max(silhouette_scores)
pam_final <- pam_models[[which.max(silhouette_scores)]]

cat(sprintf("\n*** OPTIMAL: k=%d, Silhouette=%.4f ***\n", optimal_k, optimal_sil))

# Add to dataframe
df$cluster_combined <- pam_final$clustering

# Plot
plot(pam_final, which.plots = 2, 
     main = sprintf("Combined Distance (catdist + euclidean, k=%d)", optimal_k))

# Cluster summary
cluster_summary <- df %>%
  group_by(cluster_combined) %>%
  summarise(
    n = n(),
    avg_temp = round(mean(temp_celsius), 2),
    avg_precip = round(mean(precipitation_winsorized), 2),
    avg_wind = round(mean(wind_gusts_winsorized), 2),
    avg_humidity = round(mean(humidity_percent), 2),
    pct_bad_roads = round(mean(road_conditions_bin == "1") * 100, 1),
    most_common_light = names(which.max(table(natural_light))),
    most_common_time = names(which.max(table(time_segment))),
    .groups = "drop"
  )

print(cluster_summary)
```

Now we can try different weightings:

```{r}
# Test different weight combinations
weight_combinations <- list(
  c(cat = 0.3, num = 0.7),  # Emphasize numerical
  c(cat = 0.5, num = 0.5),  # Equal weight
  c(cat = 0.7, num = 0.3)   # Emphasize categorical
)

results <- list()

for (weights in weight_combinations) {
  D_combined <- weights["cat"] * D_cat_norm + weights["num"] * D_num_norm
  pam_k <- pam(D_combined, k = optimal_k, diss = TRUE)
  
  results[[paste0("cat:", weights["cat"])]] <- list(
    silhouette = pam_k$silinfo$avg.width,
    weights = weights
  )
  
  cat(sprintf("Weights (cat=%.1f, num=%.1f): Sil=%.4f\n", 
              weights["cat"], weights["num"], 
              pam_k$silinfo$avg.width))
}

# Find best weighting
best_idx <- which.max(sapply(results, function(x) x$silhouette))
best_weights <- results[[best_idx]]$weights
cat(sprintf("\nBest weights: cat=%.1f, num=%.1f\n", 
            best_weights["cat"], best_weights["num"]))
```

Try different catdist methods:

```{r}
# Test different catdist methods for categorical variables
catdist_methods <- c("tot_var_dist", "var_mutability", "hellinger")

for (method in catdist_methods) {
  dist_cat <- catdist::cdist(df_categorical, method = method)
  D_cat <- dist_cat$distance_mat
  D_cat_norm <- D_cat / max(D_cat)
  
  D_combined <- best_weights["cat"] * D_cat_norm + best_weights["num"] * D_num_norm
  pam_k <- pam(D_combined, k = optimal_k, diss = TRUE)
  
  cat(sprintf("%s: Sil=%.4f\n", method, pam_k$silinfo$avg.width))
}
```

Compare all

```{r}
# Compare: Gower vs catdist-only vs combined
dissim_gower <- daisy(df_clean, metric = "gower")
pam_gower <- pam(dissim_gower, k = optimal_k, diss = TRUE)

pam_catdist_only <- pam(D_cat, k = optimal_k, diss = TRUE)

cat("\n=== COMPARISON ===\n")
cat(sprintf("Gower (mixed):          %.4f\n", pam_gower$silinfo$avg.width))
cat(sprintf("catdist only (3 vars):  %.4f\n", pam_catdist_only$silinfo$avg.width))
cat(sprintf("Combined (8 vars):      %.4f\n", pam_final$silinfo$avg.width))
```

We will try other weightings with different k values and hellinger method:

```{r}
# Separate variables
df_categorical <- df_clean %>% 
  select(natural_light, time_segment, road_conditions_bin)

df_numerical <- df_clean %>% 
  select(precipitation_winsorized, temp_celsius, wind_gusts_winsorized, 
         humidity_percent, dry_spells_winsorized)

# Compute distances
cat("Computing hellinger distance for categorical variables...\n")
dist_cat <- catdist::cdist(df_categorical, method = "hellinger")
D_cat <- dist_cat$distance_mat

cat("Computing euclidean distance for numerical variables...\n")
df_numerical_scaled <- scale(df_numerical)
D_num <- dist(df_numerical_scaled, method = "euclidean")
D_num <- as.matrix(D_num)

# Normalize to [0, 1]
D_cat_norm <- D_cat / max(D_cat)
D_num_norm <- D_num / max(D_num)

# Test different weight combinations
weight_cat_options <- seq(0.7, 0.9, by = 0.1)  # Test 0.7, 0.8, 0.9
k_range <- 5:10

results_matrix <- matrix(NA, nrow = length(weight_cat_options), ncol = length(k_range))
rownames(results_matrix) <- paste0("cat_", weight_cat_options)
colnames(results_matrix) <- paste0("k=", k_range)

cat("\n=== TESTING WEIGHT COMBINATIONS ===\n")
cat("(Rows = categorical weight, Columns = k)\n\n")

best_overall_sil <- -Inf
best_weight <- NULL
best_k <- NULL
best_pam <- NULL

for (i in seq_along(weight_cat_options)) {
  weight_cat <- weight_cat_options[i]
  weight_num <- 1 - weight_cat
  
  cat(sprintf("Testing weight: cat=%.1f, num=%.1f\n", weight_cat, weight_num))
  
  # Combine distances
  D_combined <- weight_cat * D_cat_norm + weight_num * D_num_norm
  
  for (j in seq_along(k_range)) {
    k <- k_range[j]
    pam_k <- pam(D_combined, k = k, diss = TRUE)
    sil <- pam_k$silinfo$avg.width
    results_matrix[i, j] <- sil
    
    cat(sprintf("  k=%d: Sil=%.4f\n", k, sil))
    
    # Track best overall
    if (sil > best_overall_sil) {
      best_overall_sil <- sil
      best_weight <- weight_cat
      best_k <- k
      best_pam <- pam_k
    }
  }
  cat("\n")
}

# Display results matrix
cat("=== RESULTS MATRIX ===\n")
print(round(results_matrix, 4))

# Find best for each weight
cat("\n=== BEST k FOR EACH WEIGHT ===\n")
for (i in seq_along(weight_cat_options)) {
  weight_cat <- weight_cat_options[i]
  best_k_for_weight <- k_range[which.max(results_matrix[i, ])]
  best_sil_for_weight <- max(results_matrix[i, ])
  cat(sprintf("cat=%.1f: k=%d, Sil=%.4f\n", 
              weight_cat, best_k_for_weight, best_sil_for_weight))
}

# Overall winner
cat(sprintf("\n*** OPTIMAL SOLUTION ***\n"))
cat(sprintf("Method: hellinger (combined)\n"))
cat(sprintf("Weights: cat=%.1f, num=%.1f\n", best_weight, 1-best_weight))
cat(sprintf("k = %d\n", best_k))
cat(sprintf("Silhouette = %.4f\n\n", best_overall_sil))

# Apply best clustering
df$cluster_hellinger_combined <- best_pam$clustering

# Visualize
par(mfrow = c(2, 2))

# 1. Heatmap of results
image(1:length(k_range), 1:length(weight_cat_options), 
      t(results_matrix), 
      xlab = "k", ylab = "Categorical Weight",
      main = "Silhouette Scores Heatmap",
      col = heat.colors(20),
      xaxt = "n", yaxt = "n")
axis(1, at = 1:length(k_range), labels = k_range)
axis(2, at = 1:length(weight_cat_options), labels = weight_cat_options)
points(which(k_range == best_k), 
       which(weight_cat_options == best_weight), 
       pch = 4, cex = 3, lwd = 3, col = "blue")

# 2. Silhouette by k for best weight
best_weight_idx <- which(weight_cat_options == best_weight)
plot(k_range, results_matrix[best_weight_idx, ], 
     type = "b", pch = 19, col = "blue", lwd = 2,
     xlab = "k", ylab = "Silhouette Width",
     main = sprintf("Best Weight (cat=%.1f)", best_weight))
points(best_k, best_overall_sil, col = "red", pch = 19, cex = 2)

# 3. Silhouette by weight for best k
best_k_idx <- which(k_range == best_k)
plot(weight_cat_options, results_matrix[, best_k_idx], 
     type = "b", pch = 19, col = "blue", lwd = 2,
     xlab = "Categorical Weight", ylab = "Silhouette Width",
     main = sprintf("Best k (k=%d)", best_k))
points(best_weight, best_overall_sil, col = "red", pch = 19, cex = 2)

# 4. Final silhouette plot
par(mfrow = c(1, 1))
plot(best_pam, which.plots = 2, 
     main = sprintf("Hellinger Combined (cat=%.1f, k=%d, Sil=%.4f)", 
                    best_weight, best_k, best_overall_sil))

# Cluster profiles
cat("\n=== CLUSTER PROFILES ===\n")
cluster_profiles <- df %>%
  group_by(cluster_hellinger_combined) %>%
  summarise(
    n = n(),
    pct = round(n()/nrow(df)*100, 1),
    # Categorical
    light = names(which.max(table(natural_light))),
    time = names(which.max(table(time_segment))),
    bad_roads_pct = round(mean(road_conditions_bin == "1")*100, 1),
    # Numerical
    temp = round(mean(temp_celsius), 1),
    precip = round(mean(precipitation_winsorized), 2),
    wind = round(mean(wind_gusts_winsorized), 1),
    humidity = round(mean(humidity_percent), 1),
    .groups = "drop"
  ) %>%
  arrange(desc(n))

print(cluster_profiles, width = Inf)

# Save best parameters for future use
cat("\n=== SAVE THESE PARAMETERS ===\n")
cat(sprintf("method = 'hellinger'\n"))
cat(sprintf("weight_categorical = %.1f\n", best_weight))
cat(sprintf("weight_numerical = %.1f\n", 1-best_weight))
cat(sprintf("optimal_k = %d\n", best_k))
```

Now we can use these parameters to investigate the composition of the clusters produced:

```{r}
library(ggplot2)

# ===== FINAL PARAMETERS =====
# method = 'hellinger'
# weight_categorical = 0.9
# weight_numerical = 0.1
# optimal_k = 10

cat("=== APPLYING FINAL CLUSTERING SOLUTION ===\n")
cat("Method: Hellinger (combined)\n")
cat("Weights: 90% categorical, 10% numerical\n")
cat("Clusters: k=10\n\n")

# Prepare data
df_categorical <- df_clean %>% 
  select(natural_light, time_segment, road_conditions_bin)

df_numerical <- df_clean %>% 
  select(precipitation_winsorized, temp_celsius, wind_gusts_winsorized, 
         humidity_percent, dry_spells_winsorized)

# Compute final distances
dist_cat <- catdist::cdist(df_categorical, method = "hellinger")
D_cat <- dist_cat$distance_mat

df_numerical_scaled <- scale(df_numerical)
D_num <- as.matrix(dist(df_numerical_scaled, method = "euclidean"))

# Normalize and combine
D_cat_norm <- D_cat / max(D_cat)
D_num_norm <- D_num / max(D_num)
D_final <- 0.9 * D_cat_norm + 0.1 * D_num_norm

# Final clustering
pam_final <- pam(D_final, k = 10, diss = TRUE)
df$cluster <- pam_final$clustering

cat(sprintf("Average Silhouette Width: %.4f\n\n", pam_final$silinfo$avg.width))

# ===== COMPREHENSIVE CLUSTER PROFILES =====
cat("=== CLUSTER PROFILES (sorted by size) ===\n")
cluster_profiles <- df %>%
  group_by(cluster) %>%
  summarise(
    n = n(),
    pct = round(n()/nrow(df)*100, 1),
    # Dominant characteristics
    light_mode = names(which.max(table(natural_light))),
    time_mode = names(which.max(table(time_segment))),
    # Light distribution
    night_pct = round(mean(natural_light == "light_night")*100, 0),
    twilight_pct = round(mean(natural_light == "light_twilight")*100, 0),
    day_pct = round(mean(natural_light == "light_day")*100, 0),
    # Road conditions
    bad_roads_pct = round(mean(road_conditions_bin == "1")*100, 1),
    # Weather
    temp_avg = round(mean(temp_celsius), 1),
    precip_avg = round(mean(precipitation_winsorized), 2),
    wind_avg = round(mean(wind_gusts_winsorized), 1),
    humidity_avg = round(mean(humidity_percent), 0),
    .groups = "drop"
  ) %>%
  arrange(desc(n))

print(cluster_profiles, width = Inf)

# ===== LABEL CLUSTERS (Give them meaningful names) =====
cat("\n=== CLUSTER LABELS ===\n")
# Based on the profiles above, you can create labels like:
cluster_labels <- cluster_profiles %>%
  mutate(
    label = paste0("C", cluster, ": ", light_mode, " + ", time_mode)
  ) %>%
  select(cluster, label, n, pct)

print(cluster_labels)

# Add labels to main dataframe
df <- df %>%
  left_join(cluster_labels %>% select(cluster, label), by = "cluster")

# ===== VISUALIZATIONS =====

# 1. Cluster sizes
p1 <- ggplot(cluster_profiles, aes(x = reorder(factor(cluster), -n), y = n)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  geom_text(aes(label = paste0(n, "\n(", pct, "%)")), vjust = -0.3, size = 3) +
  labs(title = "Accident Distribution Across 10 Clusters",
       x = "Cluster", y = "Number of Accidents") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 0))

print(p1)

# 2. Silhouette plot
plot(pam_final, which.plots = 2, 
     main = "Silhouette Plot (k=10, Hellinger Combined)")

# 3. Cluster composition by light condition
p2 <- df %>%
  group_by(cluster, natural_light) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(cluster) %>%
  mutate(pct = n/sum(n)*100) %>%
  ggplot(aes(x = factor(cluster), y = pct, fill = natural_light)) +
  geom_bar(stat = "identity") +
  labs(title = "Light Conditions by Cluster",
       x = "Cluster", y = "Percentage", fill = "Light") +
  theme_minimal() +
  scale_fill_manual(values = c("light_night" = "navy", 
                                "light_twilight" = "steelblue", 
                                "light_day" = "gold"))

print(p2)

# 4. Weather profiles by cluster
weather_long <- df %>%
  select(cluster, temp_celsius, precipitation_winsorized, 
         wind_gusts_winsorized, humidity_percent) %>%
  pivot_longer(-cluster, names_to = "variable", values_to = "value")

p3 <- ggplot(weather_long, aes(x = factor(cluster), y = value)) +
  geom_boxplot(aes(fill = factor(cluster))) +
  facet_wrap(~variable, scales = "free_y") +
  labs(title = "Weather Conditions by Cluster",
       x = "Cluster", y = "Value") +
  theme_minimal() +
  theme(legend.position = "none")

print(p3)

# ===== CLUSTER QUALITY METRICS =====
cat("\n=== CLUSTER QUALITY METRICS ===\n")
cat(sprintf("Average Silhouette Width: %.4f\n", pam_final$silinfo$avg.width))
cat("\nSilhouette Width by Cluster:\n")
print(round(pam_final$silinfo$clus.avg.widths, 4))

cat("\nCluster with lowest quality: ", 
    which.min(pam_final$silinfo$clus.avg.widths), 
    " (Sil=", round(min(pam_final$silinfo$clus.avg.widths), 4), ")\n")
cat("Cluster with highest quality: ", 
    which.max(pam_final$silinfo$clus.avg.widths), 
    " (Sil=", round(max(pam_final$silinfo$clus.avg.widths), 4), ")\n")

# Number of misclassified observations
poorly_clustered <- sum(pam_final$silinfo$widths[, "sil_width"] < 0)
cat(sprintf("\nPoorly clustered accidents: %d (%.1f%%)\n", 
            poorly_clustered, poorly_clustered/nrow(df)*100))

# ===== SAVE RESULTS =====
cat("\n=== SAVING RESULTS ===\n")

# Save the clustered data
write.csv(df, "accidents_clustered_k10.csv", row.names = FALSE)
cat("✓ Saved: accidents_clustered_k10.csv\n")

# Save cluster profiles
write.csv(cluster_profiles, "cluster_profiles_k10.csv", row.names = FALSE)
cat("✓ Saved: cluster_profiles_k10.csv\n")

# Save clustering parameters and medoids
clustering_info <- list(
  method = "hellinger_combined",
  weight_categorical = 0.9,
  weight_numerical = 0.1,
  k = 10,
  silhouette_avg = pam_final$silinfo$avg.width,
  medoids = pam_final$medoids,
  cluster_sizes = table(df$cluster)
)

saveRDS(clustering_info, "clustering_parameters_k10.rds")
cat("✓ Saved: clustering_parameters_k10.rds\n")

# Save the PAM object
saveRDS(pam_final, "pam_model_k10.rds")
cat("✓ Saved: pam_model_k10.rds\n")

cat("\n=== CLUSTERING COMPLETE! ===\n")
cat("Your 10 clusters are ready for analysis.\n")
```

```{r}
# 3. Cluster composition by light condition
p2 <- df %>%
  group_by(cluster, natural_light) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(cluster) %>%
  mutate(pct = n/sum(n)*100) %>%
  ggplot(aes(x = factor(cluster), y = pct, fill = natural_light)) +
  geom_bar(stat = "identity", color = "black") +
  labs(title = "Light Conditions by Cluster",
       x = "Cluster", y = "Percentage", fill = "Light") +
  theme_minimal() +
  scale_fill_manual(values = c("light_night" = "navy", 
                                "light_twilight" = "steelblue", 
                                "light_day" = "gold"))

print(p2)

```

```{r}
# install.packages("factoextra")
library(factoextra)
fviz_silhouette(pam_final) +
  ggplot2::labs(title = "Silhouette Plot (k = 10, Hellinger Combined)")

```

```{r}
# Create risk matrix
risk_matrix <- df %>%
  mutate(
    light_category = case_when(
      natural_light == "light_night" ~ "Night",
      natural_light == "light_twilight" ~ "Twilight",
      natural_light == "light_day" ~ "Day"
    ),
    road_category = ifelse(road_conditions_bin == "1", "Wet-Slippery", "Dry")
  ) %>%
  group_by(light_category, road_category) %>%
  summarise(n = n(), .groups = "drop") %>%
  pivot_wider(names_from = road_category, values_from = n, values_fill = 0)

cat("RISK MATRIX (Accidents by Light × Road Conditions):\n")
print(risk_matrix)
```

```{r}
# Calculate risk ratios
risk_analysis <- data.frame(
  Light = c("Day", "Twilight", "Night"),
  Bad_Roads = c(1902, 822, 803),
  Good_Roads = c(8376, 1493, 1453),
  Total = c(10278, 2315, 2256)
) %>%
  mutate(
    Pct_Bad = round(Bad_Roads / Total * 100, 1),
    Relative_Risk = round((Bad_Roads / Total) / (1902 / 10278), 2), 
    Light = factor(Light, levels = c("Day", "Twilight", "Night"))
  )

print(risk_analysis)

cat("\n=== KEY INSIGHT ===\n")
cat("Bad road conditions during:\n")
cat("- DAY:      18.5% of accidents (baseline)\n")
cat("- NIGHT:    35.6% of accidents (1.9x higher!)\n")
cat("- TWILIGHT: 35.5% of accidents (1.9x higher!)\n\n")

cat("Night/Twilight accidents are TWICE AS LIKELY to involve bad roads!\n")
```

```{r}
library(ggplot2)

# Risk by light condition
risk_plot <- risk_analysis %>%
  ggplot(aes(x = Light, y = Pct_Bad, fill = Light)) +
  geom_bar(stat = "identity", color = "black") +
  geom_hline(yintercept = 18.5, linetype = "dashed", color = "red", size = 1) +
  annotate("text", x = 3, y = 20, label = "Day baseline: 18.5%", color = "white") +
  geom_text(aes(label = paste0(Pct_Bad, "%")), vjust = -0.5, size = 5) +
  labs(title = "Wet/Slippy Road Accident Rate by Light Condition",
       subtitle = "Night and Twilight show 2x higher wet/slippy road involvement",
       x = "Light Condition", 
       y = "% of Accidents on Wet/Slippy Roads") +
  scale_fill_manual(values = c("Day" = "gold", "Twilight" = "steelblue", "Night" = "navy")) +
  theme_minimal() +
  theme(legend.position = "none") +
  ylim(0, 40)

print(risk_plot)

# Stacked bar chart
risk_matrix_long <- df %>%
  mutate(
    light_category = case_when(
      natural_light == "light_night" ~ "Night",
      natural_light == "light_twilight" ~ "Twilight",
      natural_light == "light_day" ~ "Day"
    ),
    road_category = ifelse(road_conditions_bin == "1", "Wet/Slippy Roads", "Dry Roads")
  )

p4 <- ggplot(risk_matrix_long, aes(x = light_category, fill = road_category)) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = scales::percent) +
  labs(title = "Road Condition Distribution by Light",
       x = "Light Condition", y = "Percentage", fill = "") +
  scale_fill_manual(values = c("Dry Roads" = "steelblue", "Wet/Slippy Roads" = "darkred")) +
  theme_minimal() +
  theme(legend.position = "top")

print(p4)
```

```{r}
# Add meaningful labels to your clusters
cluster_interpretations <- data.frame(
  cluster = 1:10,
  label = c(
    "Nighttime - Good Roads",                    # C1
    "Daytime - Late Morning/Midday",             # C2 (LARGEST)
    "Daytime - Afternoon",                       # C3
    "Twilight - Evening Peak",                   # C4
    "Daytime - Morning Rush (Wet)",              # C5 (2nd LARGEST)
    "Daytime - Evening Peak (Some bad roads)",   # C6
    "Daytime - Bad Roads (Wet)",                 # C7
    "Twilight - Bad Roads",                      # C8
    "Nighttime - Bad Roads (Wet)",               # C9
    "Twilight - Evening"                         # C10
  ),
  risk_profile = c(
    "Low",      # C1
    "Low",      # C2
    "Low",      # C3
    "Medium",   # C4
    "Medium",   # C5
    "Medium",   # C6
    "High",     # C7
    "High",     # C8
    "High",     # C9
    "Medium"    # C10
  )
)

print(cluster_interpretations)

# Add to your dataframe
df <- df %>%
  left_join(cluster_interpretations, by = "cluster")
```

```{r}
library(clusterSim)  # For Dunn index
library(fpc)         # For CH index and alternative calculations
library(clValid)     # Alternative package for validation

# ===== CALCULATE CLUSTER VALIDITY INDICES =====

cat("=== CLUSTER VALIDITY METRICS ===\n\n")

# Your existing distance matrix and clustering
# D_final <- your combined distance matrix
# pam_final <- your PAM clustering result with k=10

# 1. DUNN INDEX (Higher is better, range [0, ∞])
cat("Calculating Dunn Index...\n")
dunn_index <- dunn(distance = D_final, clusters = pam_final$clustering)
cat(sprintf("Dunn Index: %.4f\n", dunn_index))
cat("Interpretation: Higher = better separated clusters\n\n")

# 2. CALINSKI-HARABASZ INDEX (Higher is better)
cat("Calculating Calinski-Harabasz (CH) Index...\n")

# Need the original data (not distance matrix) for CH
# Combine scaled numerical and dummy-coded categorical
df_numerical_scaled <- scale(df_numerical)
df_categorical_dummy <- model.matrix(~ . - 1, data = df_categorical)
df_combined_for_ch <- cbind(df_numerical_scaled, df_categorical_dummy)

# Calculate CH index
ch_index <- calinhara(df_combined_for_ch, 
                      pam_final$clustering, 
                      cn = max(pam_final$clustering))
cat(sprintf("Calinski-Harabasz Index: %.2f\n", ch_index))
cat("Interpretation: Higher = better defined clusters\n\n")

# 3. DAVIES-BOULDIN INDEX (Lower is better)
cat("Calculating Davies-Bouldin Index...\n")

# Manual calculation of Davies-Bouldin
db_index <- index.DB(df_combined_for_ch, 
                     pam_final$clustering, 
                     centrotypes = "centroids")$DB
cat(sprintf("Davies-Bouldin Index: %.4f\n", db_index))
cat("Interpretation: Lower = better cluster separation\n\n")

# ===== SUMMARY TABLE =====
cat("=== SUMMARY ===\n")
validity_summary <- data.frame(
  Metric = c("Dunn Index", "Calinski-Harabasz", "Davies-Bouldin", "Silhouette"),
  Value = c(dunn_index, ch_index, db_index, pam_final$silinfo$avg.width),
  Better = c("Higher", "Higher", "Lower", "Higher"),
  Quality = c(
    ifelse(dunn_index > 0.1, "Good", "Poor"),
    ifelse(ch_index > 100, "Good", "Moderate"),
    ifelse(db_index < 1.5, "Good", ifelse(db_index < 2.5, "Moderate", "Poor")),
    ifelse(pam_final$silinfo$avg.width > 0.5, "Good", 
           ifelse(pam_final$silinfo$avg.width > 0.25, "Moderate", "Poor"))
  )
)

print(validity_summary)
```

```{r}
# ===== COMPARE METRICS FOR DIFFERENT k =====

k_range <- 4:10
results_df <- data.frame(
  k = integer(),
  Dunn = numeric(),
  CH = numeric(),
  DB = numeric(),
  Silhouette = numeric()
)

cat("\n=== TESTING k FROM 4 TO 10 ===\n")

for (k in k_range) {
  cat(sprintf("Testing k=%d...\n", k))
  
  # Cluster with current k
  pam_k <- pam(D_final, k = k, diss = TRUE)
  
  # Calculate metrics
  dunn_k <- dunn(distance = D_final, clusters = pam_k$clustering)
  ch_k <- calinhara(df_combined_for_ch, pam_k$clustering, cn = k)
  db_k <- index.DB(df_combined_for_ch, pam_k$clustering)$DB
  sil_k <- pam_k$silinfo$avg.width
  
  # Store results
  results_df <- rbind(results_df, data.frame(
    k = k,
    Dunn = dunn_k,
    CH = ch_k,
    DB = db_k,
    Silhouette = sil_k
  ))
}

print(results_df)

# Find optimal k for each metric
cat("\n=== OPTIMAL k BY EACH METRIC ===\n")
cat(sprintf("Dunn (max):       k = %d (%.4f)\n", 
            results_df$k[which.max(results_df$Dunn)], 
            max(results_df$Dunn)))
cat(sprintf("CH (max):         k = %d (%.2f)\n", 
            results_df$k[which.max(results_df$CH)], 
            max(results_df$CH)))
cat(sprintf("DB (min):         k = %d (%.4f)\n", 
            results_df$k[which.min(results_df$DB)], 
            min(results_df$DB)))
cat(sprintf("Silhouette (max): k = %d (%.4f)\n", 
            results_df$k[which.max(results_df$Silhouette)], 
            max(results_df$Silhouette)))

# Normalize and average rankings
results_normalized <- results_df %>%
  mutate(
    Dunn_norm = (Dunn - min(Dunn)) / (max(Dunn) - min(Dunn)),
    CH_norm = (CH - min(CH)) / (max(CH) - min(CH)),
    DB_norm = 1 - ((DB - min(DB)) / (max(DB) - min(DB))),  # Invert since lower is better
    Sil_norm = (Silhouette - min(Silhouette)) / (max(Silhouette) - min(Silhouette)),
    Avg_Score = (Dunn_norm + CH_norm + DB_norm + Sil_norm) / 4
  )

cat(sprintf("\nBest overall k (averaged): k = %d (score: %.3f)\n", 
            results_normalized$k[which.max(results_normalized$Avg_Score)],
            max(results_normalized$Avg_Score)))
```

```{r}
# packages
library(dplyr)

# Example column names:
# natural_light: "light_twilight", "light_night", "light_day"
# road_conditions_bin: 1 = wet/slippery, 0 = dry

df_a <- df %>%
  filter(natural_light %in% c("light_twilight", "light_night")) %>%
  mutate(is_wet = road_conditions_bin == 1)

# Counts per condition
counts <- df_a %>%
  group_by(natural_light) %>%
  summarise(wet = sum(is_wet), total = n(), .groups = "drop")

counts

# Extract counts
twilight_wet   <- counts$wet[counts$natural_light == "light_twilight"]
twilight_total <- counts$total[counts$natural_light == "light_twilight"]
night_wet      <- counts$wet[counts$natural_light == "light_night"]
night_total    <- counts$total[counts$natural_light == "light_night"]

# Proportion CIs for each
binom.test(twilight_wet, twilight_total)  # 95% CI for twilight
binom.test(night_wet,    night_total)     # 95% CI for night

# Two-proportion test (twilight vs night)
prop.test(
  x = c(twilight_wet, night_wet),
  n = c(twilight_total, night_total),
  correct = FALSE
)

```

```{r}
# install.packages("epitools")
detach("package:effectsize", unload = TRUE)
library(epitools)


# 2x2 table: rows = light (twilight/night), cols = wet vs dry
mat <- matrix(c(twilight_wet, twilight_total - twilight_wet,
                night_wet,    night_total    - night_wet),
              nrow = 2, byrow = TRUE,
              dimnames = list(c("Twilight","Night"), c("Wet","Dry")))

riskratio(mat)  # RR with CI
oddsratio(mat)  # OR with CI

```
